---
output: github_document
references:
- id: hair2018
  title: Multivariate data analysis
  author:
  - family: Hair
    given: Joseph F.
  - family: Black
    given: William C.
  - family: Babin
    given: Barry J.
  - family: Anderson
    given: Rolph E.
  edition: 8th ed.
  publisher: Cengage Learning
  type: book
  issued:
    year: 2018
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)
options(width = 100)
```
# FactorAssumptions 

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/FactorAssumptions)](https://cran.r-project.org/package=FactorAssumptions)

Set of Assumptions for Factor and Principal Component Analysis.
Tests for Kaiser-Meyer-Olkin (KMO) and Communalities in a dataset. It provides a final sample by removing variables in a iterable manner while keeping account of the variables that were removed in each step.

## Installation

* You can install from CRAN
* You can install the released version of FactorAssumptions from [Github](https://github.com/storopoli/FactorAssumptions)

```{r install, eval=FALSE}
install.packages("FactorAssumptions")

## install devtools if not already
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
## install FactorAssumptions from Github
devtools::install_github("storopoli/FactorAssumptions")
```

## What is KMO and Communalities?

*Factor Analysis* and *Principal Components Analysis* (PCA) have some precautions and assumptions to be observed (@hair2018). 

The first one is the KMO (Kaiser-Meyer-Olkin) measure, which measures the proportion of variance among the variables that can be derived from the common variance, also called systematic variance. KMO is computed between 0 and 1. Low values (close to 0) indicate that there are large partial correlations in comparison to the sum of the correlations, that is, there is a predominance of correlations of the variables that are problematic for the factorial/principal component analysis. @hair2018 suggest that individual KMOs smaller than 0.5 be removed from the factorial/principal component analysis. Consequently, this removal causes the overall KMO of the remaining variables of the factor/principal component analysis to be greater than 0.5.

The second assumption of a valid factor or PCA analysis is the communality of the rotated variables. The commonalities indicate the common variance shared by factors/components with certain variables. Greater communality indicated that a greater amount of variance in the variable was extracted by the factorial/principal component solution. For a better measurement of factorial/principal component analysis, communalities should be 0.5 or greater (@hair2018).

## Vignette

I encourage you to check the vignette on how to use the package.

# References
