---
title: "How to use FactorAssumptions"
author: "Jose Eduardo Storopoli"
date: "7/16/2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use FactorAssumptions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{FactorAssumptions}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## FactorAssumptions

Set of Assumptions for Factor and Principal Component Analysis

Description:Tests for Kaiser-Meyer-Olkin (KMO) and Communalities in a dataset. It provides a final sample by removing variables in a iterable manner while keeping account of the variables that were removed in each step.

First we will load a sample dataset `bfi` from `psych` and load the package `FactorAssumptions`

```{r bfi}
library(FactorAssumptions, quietly = T, verbose = F)
bfi_data <- bfi
#Remove rows with missing values and keep only complete cases
bfi_data <- bfi_data[complete.cases(bfi_data),]
head(bfi_data)
```

## Performing the KMO Assumptions

First we will perform the $KMO > 0.5 assumption$ for all individuals variables in the dataset with the `kmo_optimal_solution` function

```{r KMO}
kmo_bfi <- kmo_optimal_solution(bfi_data)
```

Note that the `kmo_optimal_solution` outputs a list:

1. the final solution as `df`
2. removed variables with $invidual KMO < 0.5$ as `removed`
3. Anti-image covariance matrix as `AIS`
4. Anti-image correlation matrix as `AIR`

In our case none of the variables were removed due to low individual KMO values

```{r removed_kmo}
kmo_bfi$removed
```

## Determining how many factors to extract

Now let's use the function `how_many_factors` to determine how many factors to retain based on:

* Eigenvalues
* Parallel Analysis
* Optimal Coordinates
* Acceleration Factor

*Note*: it will generate a 1,000 replications as default for the Paralle Analysis, so maybe it will take a while to run.

```{r how_many_factors}
how_many_factors(bfi_data)
```

## Performing the Communalities Assumptions

Since the output of the `how_many_factors` suggest seven factors we will then perform the assumptions for all $individual communalities > 0.5$ with the argument `nfactors` set to 7.

*Note*: we are using the `df` generated from the `kmo_optimal_solution` function
*Note 2*: the default of rotation employed by the `communalities_optimal_solution` is `varimax`

```{r communalities}
comm_bfi <- communalities_optimal_solution(kmo_bfi$df, nfactors = 7)
```

Note that the `communalities_optimal_solution` outputs a list:

1. the final solution as `df`
2. removed variables with $invidual communalities < 0.5$ as `removed`
3. A table with the communalities loadings from the variables final iteration as `loadings`
4. Results of the final iteration of the `principal` function from `psych` package as `pca_results`

In our case 3 variables were removed in an iterable fashion due to low individual communality values. And they are listed from the lowest communality that were removed until rendered an optimal solution.

```{r removed_comm}
comm_bfi$removed
```

And finally we arrive at our final principal components analysis rotated matrix. You can export it as a csv with `write.csv` or `write.csv2`

```{r final_solution}
comm_bfi$pca_results
```
